### 2021-07-31 response.tex

The exact shape of the distribution depends on several factors---including the shape of latent and infectious periods \citep{lloyd2001realistic,wearing2005appropriate,roberts2007model} as well as more detailed life history of a disease \citep{huber2016quantitative}---and thus can be difficult to estimate.
While it is possible to consider general forms of generation-interval distributions \citep{miller2010epidemics,svensson2015influence}, summarizing the distribution in terms of its mean and variability---for example, by using the Gamma distribution---can still provide a robust link between epidemic speed and strength for real diseases and yield important biological insights \citep{park2019practical}.
In particular, several studies have noted, in many contexts, that mechanisms that increase the mean generation interval increase epidemic strength $\RR$ given observed epidemic speed $r$ \citep{eaton2014proportion,powers2014impact,weitz2015modeling,park2020time}.''

Assuming Gamma-distributed generation intervals is different from the standard Gamma-chain trick. We now explain this later in Section 2:

``Various Gamma-generation-interval assumptions have been widely used in epidemic modeling, including for models of SARS-CoV-2 \citep{doi:10.1098/rsif.2020.0144}.
The Gamma-generation-interval assumption includes as a special case models that assume exponentially distributed generation intervals (when $\kappa=1$), corresponding to the SIR model \citep{anderson1991infectious}.
We note that when the infectious periods are Gamma distributed---another standard assumption in epidemic modeling---the resulting generation interval does not follow the Gamma distribution;
in particular, the mean generation interval can be shorter than the mean infectious period because transmission occurs throughout the infectious period until recovery (see \cite{roberts2007model} for detailed discussion).''

@@

\revtext{Other comments:}

\revtext{Topic on which there is a considerable confusion in terminology in the literature.  In addition to "generation interval", terms "generation time", "transmission interval", "serial interval", and others have been used in the literature, all with related by different definitions.  This manuscripts definition "time between infection and transmission" implies a single transmission event, but Kenah pointed out that time-contraction appears when allowing multiple transmission events (see Champredon 2015).  A clearer definition up front will help readers.  Fortunately, the results don't depend on the specific definition.}

Thank you for the suggestions. We have tried to clarify what we mean by the generation interval. In particular, often-mentioned time contractions of generation intervals apply to the forward realized generation interval, which measure time between actual infection events, rather than the intrinsic generation interval, which measure the intrinsic infectiousness of infected individuals. Under strength-like interventions, the intrinsic generation interval does not vary across time, even though the realized generation interval contracts. These distinctions are explained in Champredon 2015, and we have tried to clarify these ideas again here in Section 2:

``Under constant-strength changes that reduce the transmission rate by a constant amount regardless of the age of infection, the instantaneous generation-interval distribution does not change across time \citep{fraser2007estimating}.
In this case, the instantaneous generation-interval distribution is also often referred to as the intrinsic generation-interval distribution \citep{champredon2015intrinsic,champredon2018two,gostic2020practical,park2020time}---
for example, the standard SEIR model can be equivalently expressed as a renewal equation with time-invariant intrinsic generation-interval distribution $g(\tau)$ as shown in \citep{champredon2018equivalence}.
The instantaneous generation-interval distribution is also different from the realized generation-interval distribution, which measures time between actual infection events (see \citep{champredon2015intrinsic}).
Previous studies have noted, in many contexts, that the realized generation intervals can contract due to susceptible depletion---a special case of constant-strength changes \citep{kenah2008generation,nishiura2010time,champredon2015intrinsic}---even though the instantaneous generation-interval distribution remains unchanged in this scenario.
While the instantaneous generation-interval distribution can change under speed-like changes (see \cite{fraser2007estimating} and Section 6--7 for detailed discussions), assuming a time-invariant instantaneous generation-interval distribution is often appropriate in the context of SARS-CoV-2, given that the dynamics of its spread have been primarily driven by strength-like intervention, such as social distancing measures \citep{flaxman2020Rt} and vaccination \citep{moore2021vaccination}.
Indeed, many dynamical models of SARS-CoV-2 infections have solely relied on constant-strength changes, either implicitly or explicitly assuming a time-invariant intrinsic generation-interval distribution (e.g., \citep{flaxman2020Rt,gostic2020practical,brauner2021inferring}).
Therefore, we neglect changes in the intrinsic generation-interval distribution over time for now and focus on the impact of constant-strength changes on the inference of dynamics of new SARS-CoV-2 variants.
We revisit these ideas in Section 6 and compare the effects of constant-speed interventions with those of constant-strength interventions.''

@@

We further note that a fixed generation-interval distribution does not rely on the assumption that the rate of depletion of susceptibles over time is negligible.
This is due to an ongoing confusion between the intrinsic and realized generation interval as we explained above.
For example, as shown in Champredon et al 2015 and 2018, the standard SEIR model can be equivalently written in terms of the renewal equation using a fixed \emph{intrinsic} generation-interval distribution, and we can still observe the contraction of the \emph{realized} generation interval for this model.
We have added a paragraph in Section 2 to clarify this distinction (please see response to reviewer 1 for text).

\revtext{(page 2, lines 55–60) It would help to clarify what is meant by “strengthlike” and “speed-like” in interventions. From the examples, it appears that
speed-like interventions work by identifying potential infections quickly
(effectively ending the infectious period early while not necessarily altering
the hazard of infectious contact) and strength-like interventions work by
reducing the hazard of infectious contact (while not necessarily altering
the duration of infectiousness). A good explanation of these ideas is on
lines 228–231 on page 10.}

We now define them more clearly throughout the text, including in the introduction:

``For example, when epidemic strength $\RR$ is fixed, assuming longer generation intervals leads to a slower epidemic growth (lower $r$), making the epidemic look easier to control with a constant-speed intervention, which isolates infected individuals at a constant hazard and therefore affects late transmission disproportionately.
As shown in \cite{doi:10.1098/rspb.2020.1556}, the hazard of isolation must be greater than the epidemic speed $r$ to reach the $r=0$ threshold.
In this same case, however, the uncertainty in the generation-interval distribution is expected to have no effects on the effectiveness of a constant-strength intervention, which reduces the hazard of infectious contacts and therefore affects transmission by an equal amount regardless of age of infection.
Conversely, when epidemic speed $r$ is fixed, assuming longer generation intervals leads to a higher estimate of epidemic strength ($\RR$), making the epidemic look harder to control with a constant-strength interventions.
Constant strength and speed are idealized representations of real-life interventions, which can be either strength-like (e.g., vaccination and social distancing) or speed-like (e.g., contact tracing and isolation) depending on how their effectiveness varies across the generation interval (see \cite{doi:10.1098/rspb.2020.1556} and Discussion).''

### 2021-08-03 newvariant.tex
done for now?

\usepackage{natbib}
\setcitestyle{numbers} 
\setcitestyle{square}
\bibliographystyle{prsb}
\date{\today}

@@

Inferring the relative strength (i.e., the ratio of reproduction numbers) and relative speed (i.e., the difference between growth rates) of new SARS-CoV-2 variants compared to their wild types is critical to predicting and controlling the course of the current pandemic.
We explore how differences in the generation-interval distributions affect the inference of relative strength and speed in light of recent evidence that the Alpha and Delta variants may have different infectiousness profiles (and therefore, the generation interval) from prior dominant lineages.
Neglecting these differences gives biased estimates of relative strength, which can vary systematically with changes in the incidence of infections caused by the variant.
We further argue that holding relative strength constant is generally more appropriate in the context of SARS-CoV-2 and illustrate how relative strength and speed are affected by two idealized interventions.
A constant-strength intervention, which reduces transmission rates by a constant amount, affects relative speed but not relative strength---this is analogous to vaccination and social distancing.
A constant-speed intervention, which isolates infected individuals at a constant hazard, affects relative strength but not relative speed---this is analogous to contact tracing.
Changes in the generation-interval distribution can alter the effectiveness of these interventions.

@@

Estimating variant epidemic strength and speed remains a key question in understanding the threat of SARS-CoV-2 variants of concern (VoCs) \citep{switzerland2021variant, davies2021estimated, di2021impact,graham2021changes, leung2021early, volz2021transmission,zhao2021,campbell2021increased}.
Epidemic ``strength'' is measured by the reproduction number $\RR$---a unitless quantity representing the average number of new infections caused by a typical infection. 
Disease can spread in a population if $\RR>1$ \citep{diekmann1990definition}.
The epidemic strength also determines the final size of an epidemic in a homogeneous population \citep{anderson1991infectious}.
Epidemic ``speed'' is characterized by the growth rate $r$, which has units of $1/\mathrm{time}$ and describes the exponential rate of disease spreads at the population level.
Like epidemic strength, epidemic speed also determines conditions for disease elimination: $r=0$ is a threshold equivalent to $\RR=1$ under constant conditions.

Epidemic speed can be typically estimated from incidence of infection during the exponential growth period \citep{mills2004transmissibility,nishiura2009transmission,ma2014estimating}, but epidemic strength is difficult to measure directly.
Instead, epidemic strength is often inferred from the observed epidemic speed using the generation-interval distribution, an approach popularized by \citep{wallinga2007generation}.
The generation interval is defined as the time between infection and transmission and provides information about the time scale of individual-level transmission \citep{svensson2007note}.
The generation interval is also distinctly different from other so-called ``transmission intervals'' that measure time between successive infections---this includes the serial interval, which is defined as the time between symptom onsets in an infector-infectee pair \citep{fine2003interval,grassly2008mathematical,britton2019estimation,ali2020serial,park2021forward}.

@@

Analyses of new variants have typically characterized \emph{relative} strength (i.e., the ratio of reproduction numbers $\Rv/\Rw$, here called $\rho$) and speed (i.e., the difference between growth rates $\rv-\rw$, here called $\delta$) of the variants.
Many studies have tried to estimate the relative strength of variants from the observed relative speed \citep{davies2021estimated, leung2021early, volz2021transmission,zhao2021,campbell2021increased}.
Instead (or additionally), some studies have assumed a value of the relative strength and tried to predict its relative speed \citep{davies2021estimated,di2021impact}.
While both approaches are reasonable, holding different quantities constant (i.e., strength or speed) can lead to different conclusions about the spread of the disease and its control because two quantities are linked by the generation-interval distribution \citep{doi:10.1098/rspb.2020.1556}.

For example, when epidemic strength $\RR$ is fixed, assuming longer generation intervals leads to a slower epidemic growth (lower $r$), making the epidemic look easier to control with a constant-speed intervention, which isolates infected individuals at a constant hazard and therefore affects late transmission disproportionately.
As shown in \cite{doi:10.1098/rspb.2020.1556}, the hazard of isolation must be greater than the epidemic speed $r$ to reach the $r=0$ threshold.
In this same case, however, the uncertainty in the generation-interval distribution is expected to have no effects on the effectiveness of a constant-strength intervention, which reduces the hazard of infectious contacts and therefore affects transmission by an equal amount regardless of age of infection.
Conversely, when epidemic speed $r$ is fixed, assuming longer generation intervals leads to a higher estimate of epidemic strength ($\RR$), making the epidemic look harder to control with a constant-strength interventions.
Constant strength and speed are idealized representations of real-life interventions, which can be either strength-like (e.g., vaccination and social distancing) or speed-like (e.g., contact tracing and isolation) depending on how their effectiveness varies across the generation interval (see \cite{doi:10.1098/rspb.2020.1556} and Discussion).

Most studies have assumed that the previous dominant strain (wild type) and variants have identical generation-interval distributions, but Kissler \textit{et al.} \citep{kissler2021densely} suggested that the Alpha variant may have a longer duration of infection: 13.3 days (90\% CI: 10.1--16.5) for the new variant and 8.2 days (90\% CI: 6.5--9.7) for the wild type.
Longer duration of infection suggests that the mean generation interval of the Alpha variant is likely to be longer than that of the wild type \citep{lloyd2001realistic,wearing2005appropriate,roberts2007model}.
In contrast, some studies have suggested that the more recent Delta variant may have a shorter generation interval \citep{li2021viral, zhang2021transmission}.
Modeling studies have also considered the possibility that the faster growth rate of new variants may be driven, in part, by shorter generation intervals \citep{davies2021estimated,volz2021transmission}.
However, linking strength and speed is complicated given that the generation-interval distribution depend on many factors including behavior: 
for example, self-isolation after symptom onset will lead to shorter generation intervals.

@@

Neglecting the (relatively slow) rate of new mutations and assuming homogeneous mixing, the current incidence of infection $i_x(t)$ caused by each strain $x$---either the wild type (``wt'') or the variant (``var'')---can be expressed in terms of their previous incidence $i_x(t-\tau)$ and the rate at which secondary infections are generated at time $t$ by individuals infected $\tau$ time units ago $K_x(t, \tau)$:
\begin{equation}
i_x(t) = \int_0^\infty i_x(t-\tau) K_x(t, \tau) \dtau.
\end{equation}
This framework provides a flexible way of modeling disease dynamics and generalizes a wide range of compartmental models, including the SEIR model \citep{heesterbeek1996concept, diekmann2000mathematical, roberts2004modelling, aldis2005integral,breda2012formulation, champredon2018equivalence}.

The integral of the kernel is referred to as the instantaneous reproduction number:
\begin{equation}
\RR_x(t) = \int K_x(t, \tau) \dtau
\label{eq:instR}
\end{equation}
which is defined as secondary infections that would be caused by an individual infected at time $t$ if conditions were to remain the same \citep{fraser2007estimating}.
The instantaneous reproduction number is a particular kind of weighted average of infectiousness of previously infected individuals at time $t$---in particular, it is weighted by the total relative infectiousness at time $t$, rather than by the actual number of infected individuals present.
The normalized kernel $g_x(t, \tau) = K_x(t, \tau)/\RR_x(t)$---which we refer to as the instantaneous generation-interval distribution---describes the relative contribution of previously infected individuals to current incidence $i_x(t)$ and provides information about the time scale of disease transmission.
Like the instantaneous reproduction number, the instantaneous generation-interval distribution describes when an individual infected at time $t$ will transmit infections if conditions were to remain the same.
Both the instantaneous reproduction number and the generation-interval distribution depend on many factors, including intrinsic infectiousness of an infected individual, non-pharmaceutical interventions, awareness-driven behavior, and population-level susceptibility \citep{fraser2007estimating}.

Under constant-strength changes that reduce the transmission rate by a constant amount regardless of the age of infection, the instantaneous generation-interval distribution does not change across time \citep{fraser2007estimating}.
In this case, the instantaneous generation-interval distribution is also often referred to as the intrinsic generation-interval distribution \citep{champredon2015intrinsic,champredon2018two,gostic2020practical,park2020time}---
for example, the standard SEIR model can be equivalently expressed as a renewal equation with time-invariant intrinsic generation-interval distribution $g(\tau)$ as shown in \citep{champredon2018equivalence}.
The instantaneous generation-interval distribution is also different from the realized generation-interval distribution, which measures time between actual infection events (see \citep{champredon2015intrinsic}).
Previous studies have noted, in many contexts, that the realized generation intervals can contract due to susceptible depletion---a special case of constant-strength changes \citep{kenah2008generation,nishiura2010time,champredon2015intrinsic}---even though the instantaneous generation-interval distribution remains unchanged in this scenario.
While the instantaneous generation-interval distribution can change under speed-like changes (see \cite{fraser2007estimating} and Section 6--7 for detailed discussions), assuming a time-invariant instantaneous generation-interval distribution is often appropriate in the context of SARS-CoV-2, given that the dynamics of its spread have been primarily driven by strength-like intervention, such as social distancing measures \citep{flaxman2020Rt} and vaccination \citep{moore2021vaccination}.
Indeed, many dynamical models of SARS-CoV-2 infections have solely relied on constant-strength changes, either implicitly or explicitly assuming a time-invariant intrinsic generation-interval distribution (e.g., \citep{flaxman2020Rt,gostic2020practical,brauner2021inferring}).
Therefore, we neglect changes in the intrinsic generation-interval distribution over time for now and focus on the impact of constant-strength changes on the inference of dynamics of new SARS-CoV-2 variants.
We revisit these ideas in Section 6 and compare the effects of constant-speed interventions with those of constant-strength interventions.

Over a short period of time, we can assume that epidemiological conditions remain roughly constant: $\RR_x(t) \approx \RR_x$ and $g_x(t, \tau) \approx g_x(\tau)$, in which case the incidence of infections changes exponentially.
Here, we use the term ``epidemiological conditions'' to broadly refer to all factors that affect transmission---mathematically, they are captured by the kernel $K(t, \tau)$.
In the context of SARS-CoV-2 infections, we are essentially assuming that the changes in susceptible pool, behavior, and contact rates are usually small over a short period of time.
Then, the incidence of each strain grows (or decays) exponentially at rate $r_x$, satisfying the Euler-Lotka equation \citep{wallinga2007generation}:
\begin{equation}
\frac{1}{\RR_x} = \int_0^\infty \exp(- r_x \tau) g_x(\tau) \dtau.
\end{equation}
We can approximate this $r$--$\RR$ relationship by assuming that the generation-interval distribution is Gamma-distributed, and summarizing it using the mean generation interval $\bar{G}_x$ and the squared coefficient of variation $\kappa_x$ \citep{park2019practical}:
\begin{equation}
\RR_x \approx (1 + \kappa_x r_x \bar{G}_x)^{1/\kappa_x}.
\end{equation}
Various Gamma-generation-interval assumptions have been widely used in epidemic modeling, including for models of SARS-CoV-2 \citep{doi:10.1098/rsif.2020.0144}.
The Gamma-generation-interval assumption includes as a special case models that assume exponentially distributed generation intervals (when $\kappa=1$), corresponding to the SIR model \citep{anderson1991infectious}.
We note that when the infectious periods are Gamma distributed---another standard assumption in epidemic modeling---the resulting generation interval does not follow the Gamma distribution;
in particular, the mean generation interval can be shorter than the mean infectious period because transmission occurs throughout the infectious period until recovery (see \cite{roberts2007model} for detailed discussion).

We use this framework to investigate how inferences about strength and speed of the variant depend on our assumptions about the underlying generation-interval distributions.
Here, we focus on differences in mean generation intervals, assuming that both the squared coefficient of variation and distribution shape of the variant are otherwise the same as for the wild type: $\kappa_{\mathrm{wt}} = \kappa_{\mathrm{var}} = \kappa$.
Changes in shape can be important \citep{miller2010epidemics,svensson2015influence}, and we do not investigate them here. 
We do note that we expect a distribution with higher coefficient of variation to allow for more early transmission, and thus to have qualitatively similar effects to a distribution with a shorter mean in many cases \citep{park2019practical}.

@@

While epidemic speed $r_x$ can often be estimated directly from incidence of infection during the exponential growth period \citep{mills2004transmissibility,nishiura2009transmission,ma2014estimating},
studies of new SARS-CoV-2 variants have mostly focused on characterizing changes in the \emph{proportion} of a new variant \citep{switzerland2021variant, davies2021estimated, di2021impact, graham2021changes, leung2021early, volz2021transmission,zhao2021}.
Focusing on proportions has the advantage, because changes in proportions are less sensitive to changes in testing and to other transient effects that would affect variants and wild type viruses similarly.
When incidence is changing exponentially ($i_x(t) = i_x(t_0) \exp(r_x t)$), the proportion of the new variant $p(t)$ follows a logistic growth curve \citep{switzerland2021variant,davies2021estimated}:
\begin{align}
p(t) &= \frac{\iv(t_0) \exp(\rv t)}{\iw(t_0) \exp(\rw t) + \iv(t_0) \exp(\rv t)},
\\ &= \frac{1}{1 + \left(\iw(t_0)/\iv(t_0)\right) \exp(-\delta t)},
\end{align}
where the relative speed of the variant $\delta = \rv - \rw$ can be estimated as the slope of the log odds of $p$  vs.~time.
When more than two strains are co-circulating, the picture is more complicated \citep{campbell2021increased}; we focus here on comparing two strains at a time.

\begin{figure}[!t]
\includegraphics[width=\textwidth]{relstrength.pdf}
\caption{
\textbf{Relative strength of the new variant assuming a fixed speed advantage $\delta$ under five epidemiological conditions.}
The relative strength of the new variant $\rho$ conditional on the speed of the wild type $\rw$; the ratio between the mean generation interval of the new variant $\Gv$ and that of the wild type $\Gw$; and the squared coefficient of variation in generation intervals $\kappa$.
The relative strength of the new variant $\rho$ is calculated using $\delta=0.1\pday$, $\Gw = 5\days$, and $\kappa = 1/5$.
Assumed values of $\rw$ (and therefore $\rv$) are shown in the top right corners of each panel.
}
\label{fig:relstrength}
\end{figure}

@@

We start by taking the mean generation interval of the wild type to be $\Gw = 5\days$ \citep{ferretti2020quantifying}; and the squared coefficient of variation of generation intervals to be $\kappa=0.2$ \citep{ferretti2020quantifying} for both the variant and the wild type.
As noted above, we assume throughout that the variant and the wild type can be approximated with gamma distributions with equal $\kappa$, and only consider differences in the mean.
We evaluate the estimates of relative strength $\rho$ across a wide range of $\kappa$ from 0 (fixed-length generation intervals) to 1 (exponential distribution).
To further explore how inference depends on underlying epidemiological conditions, we consider five scenarios: (1) $\rw < \rv < 0$, (2) $\rw < \rv = 0$, (3) $\rw < 0 < \rv$, (4) $0 = \rw < \rv$, and (5) $0 < \rw < \rv$.
In all of these cases, we assume $\delta = \rv - \rw = 0.1\pday$ based on the observed relative growth rate of the Alpha variant in the UK \citep{davies2021estimated}.

>>>>

\usepackage[sort&compress]{natbib}
\setcitestyle{numbers} 
\setcitestyle{square}
\bibliographystyle{prsb}
\date{\today}

@@

Inferring the relative strength (i.e., the ratio of reproduction numbers) and relative speed (i.e., the difference between growth rates) of new SARS-CoV-2 variants compared to the wild types, from which they emerged, is critical to predicting and controlling the course of the current pandemic.
We explore how differences in the generation-interval distributions affect the inference of relative strength and speed in light of recent evidence that the Alpha and Delta variants may have different generation intervals from prior dominant lineages.
Neglecting these differences gives biased estimates of relative strength, which can vary systematically with changes in the incidence of infections caused by the variant.
We further argue that holding relative strength constant is generally more appropriate for estimating the transmission advantage of new SARS-CoV-2 variants by comparing how relative strength and speed change over time under two idealized interventions of constant strength and speed.
Constant-strength interventions like idealized vaccination and social distancing, which reduce transmission rates by a constant amount, affect relative speed but not relative strength.
Constant-speed interventions like idealized contact tracing, which isolate infected individuals at a constant hazard, affect relative strength but not relative speed.
Changes in the generation-interval distribution can alter the effectiveness of these interventions.

@@

Estimating variant epidemic strength and speed remains a key question in understanding the threat of SARS-CoV-2 variants of concern (VoCs) \cite{switzerland2021variant,davies2021estimated,di2021impact,graham2021changes,leung2021early,volz2021transmission,zhao2021,campbell2021increased}.
Epidemic ``strength'' is measured by the reproduction number $\RR$---a unitless quantity representing the average number of new infections caused by a typical infection. 
Disease can spread in a population if $\RR>1$ \citep{diekmann1990definition}.
The epidemic strength also determines the final size of an epidemic in a homogeneous population \citep{anderson1991infectious}.
Epidemic ``speed'' is characterized by the growth rate $r$, which has units of $1/\mathrm{time}$ and describes the exponential rate of disease spread at the population level.
Like epidemic strength, epidemic speed also determines conditions for disease elimination: $r=0$ is a threshold equivalent to $\RR=1$ under constant conditions.

Epidemic speed can be typically estimated from time series of incidence of infection during the exponential growth period \citep{mills2004transmissibility,nishiura2009transmission,ma2014estimating}, but epidemic strength is difficult to measure directly.
Instead, epidemic strength is often inferred from the observed epidemic speed using the generation-interval distribution, an approach popularized by \citep{wallinga2007generation}.
The generation interval, defined as the time between infection and transmission, provides information about the time scale of individual-level transmission \citep{svensson2007note}.
The generation interval is also distinctly different from other so-called ``transmission intervals'' that measure time between successive infections---including the serial interval, which is defined as the time between symptom onsets in an infector-infectee pair \citep{fine2003interval,grassly2008mathematical,britton2019estimation,ali2020serial,park2021forward}.

@@

Analyses of new variants have typically characterized \emph{relative} strength (i.e., the ratio of reproduction numbers $\Rv/\Rw$, here called $\rho$) and speed (i.e., the difference between growth rates $\rv-\rw$, here called $\delta$) of the variants.
Many studies have tried to estimate the relative strength of variants from the observed relative speed \citep{davies2021estimated, leung2021early, volz2021transmission,zhao2021,campbell2021increased}.
Instead (or additionally), some studies have assumed a value of the relative strength and tried to predict its relative speed \citep{davies2021estimated,di2021impact}.
While both approaches are reasonable, holding different quantities constant (i.e., strength or speed) can lead to different conclusions about the spread of the disease and its control because the two quantities are linked by the generation-interval distribution \citep{doi:10.1098/rspb.2020.1556}.

For example, when epidemic strength $\RR$ is fixed, assuming longer generation intervals leads to a slower estimated epidemic growth rate (lower $r$), making the epidemic look easier to control with a constant-speed intervention, which isolates infected individuals at a constant hazard and therefore affects late transmission disproportionately.
As shown in \cite{doi:10.1098/rspb.2020.1556}, the hazard of isolation must be greater than the epidemic speed $r$ to reach the $r=0$ threshold.
In this same case, however, variation in the generation-interval distribution has no effects on the effectiveness of a constant-strength intervention, which reduces the hazard of infectious contacts and therefore affects transmission by an equal amount regardless of age of infection.
Conversely, when epidemic speed $r$ is fixed, assuming longer generation intervals leads to a higher estimate of epidemic strength ($\RR$), making the epidemic look harder to control with a constant-strength interventions.
Constant strength and speed are idealized representations of real-life interventions, which can be either strength-like (e.g., vaccination and social distancing) or speed-like (e.g., contact tracing and isolation) depending on how their effectiveness varies across the generation interval (see \cite{doi:10.1098/rspb.2020.1556} and Discussion).

Most studies have assumed that the previous dominant strain (wild type) and variants have identical generation-interval distributions, but Kissler \textit{et al.} \citep{kissler2021densely} suggested that the Alpha variant may have a longer duration of infection: 13.3 days (90\% CI: 10.1--16.5) for the new variant and 8.2 days (90\% CI: 6.5--9.7) for the wild type.
Longer duration of infection suggests that the mean generation interval of the Alpha variant is likely to be longer than that of the wild type \citep{lloyd2001realistic,wearing2005appropriate,roberts2007model}.
In contrast, some studies have suggested that the more recent Delta variant may have a shorter generation interval \citep{li2021viral,zhang2021transmission}.
Modeling studies have also considered the possibility that the faster growth rate of new variants may be driven, in part, by shorter generation intervals \citep{davies2021estimated,volz2021transmission}.
However, linking strength and speed is complicated given that the generation-interval distribution depends on many factors including behavior: 
for example, self-isolation after symptom onset will lead to shorter generation intervals.

@@

Neglecting the (relatively slow) rate of new mutations and assuming homogeneous mixing, the current incidence of infection $i_x(t)$ caused by each strain $x$---either the wild type (``wt'') or the variant (``var'')---can be expressed in terms of their previous incidence $i_x(t-\tau)$ and the rate $K_x(t, \tau)$ at which secondary infections are generated at time $t$ by individuals infected $\tau$ time units ago:
\begin{equation}
i_x(t) = \int_0^\infty i_x(t-\tau) K_x(t, \tau) \dtau.
\end{equation}
This framework provides a flexible way of modeling disease dynamics and generalizes a wide range of compartmental models, including the SEIR model \citep{heesterbeek1996concept, diekmann2000mathematical, roberts2004modelling, aldis2005integral,breda2012formulation, champredon2018equivalence}.

The integral of the kernel is referred to as the instantaneous reproduction number:
\begin{equation}
\RR_x(t) = \int K_x(t, \tau) \dtau
\label{eq:instR}
\end{equation}
which is defined as the expected number of secondary infections that would be caused by an individual infected at time $t$ if conditions were to remain the same \citep{fraser2007estimating}.
The instantaneous reproduction number is a particular kind of weighted average of infectiousness of previously infected individuals at time $t$---in particular, it is weighted by the total relative infectiousness at time $t$, rather than by the actual number of infected individuals present.
The normalized kernel $g_x(t, \tau) = K_x(t, \tau)/\RR_x(t)$---which we refer to as the instantaneous generation-interval distribution---describes the relative contribution of previously infected individuals to current incidence $i_x(t)$ and provides information about the time scale of disease transmission.
Like the instantaneous reproduction number, the instantaneous generation-interval distribution describes when an individual infected at time $t$ will transmit infections if conditions were to remain the same.
Both the instantaneous reproduction number and the generation-interval distribution depend on many factors, including intrinsic infectiousness of an infected individual, non-pharmaceutical interventions, awareness-driven behavior, and population-level susceptibility \citep{fraser2007estimating}.

Under constant-strength changes that reduce the transmission rate by a constant amount regardless of the age of infection, the instantaneous generation-interval distribution does not change through time \citep{fraser2007estimating}.
In this case, the instantaneous generation-interval distribution is also often referred to as the intrinsic generation-interval distribution \citep{champredon2015intrinsic,champredon2018two,gostic2020practical,park2020time}---
for example, the standard SEIR model can be equivalently expressed as a renewal equation with time-invariant intrinsic generation-interval distribution $g(\tau)$ as shown in \citep{champredon2018equivalence}.
The instantaneous generation-interval distribution is also different from the realized generation-interval distribution, which measures time between actual infection events \citep{champredon2015intrinsic}.
Previous studies have noted, in many contexts, that the realized generation intervals can contract due to susceptible depletion---a special case of constant-strength changes \citep{kenah2008generation,nishiura2010time,champredon2015intrinsic}---even though the instantaneous generation-interval distribution remains unchanged in this scenario.
While the instantaneous generation-interval distribution can change under speed-like changes (see \cite{fraser2007estimating} and Section 6--7 for detailed discussions), assuming a time-invariant instantaneous generation-interval distribution is often appropriate in the context of SARS-CoV-2, given that control strategies against its spread have been primarily strength-like, including social distancing measures \citep{flaxman2020Rt} and vaccination \citep{moore2021vaccination}.
Indeed, many dynamical models of SARS-CoV-2 infections have solely relied on constant-strength changes, either implicitly or explicitly assuming a time-invariant intrinsic generation-interval distribution (e.g., \citep{flaxman2020Rt,gostic2020practical,brauner2021inferring}).
Therefore, we neglect changes in the intrinsic generation-interval distribution over time for now and focus on the impact of constant-strength changes on the inference of dynamics of new SARS-CoV-2 variants.
We revisit these ideas in Section 6 and compare the effects of constant-speed interventions with those of constant-strength interventions.

Over a short period of time, we can assume that epidemiological conditions remain roughly constant: $\RR_x(t) \approx \RR_x$ and $g_x(t, \tau) \approx g_x(\tau)$, in which case the incidence of infections changes exponentially.
Here, we use the term ``epidemiological conditions'' to broadly refer to all factors that affect transmission---mathematically, they are captured by the kernel $K(t, \tau)$.
In the context of SARS-CoV-2 infections, we are essentially assuming that the changes in susceptible pool, behavior, and contact rates are usually small over a short period of time.
Then, the incidence of each strain grows (or decays) exponentially at rate $r_x$, satisfying the Euler-Lotka equation \citep{wallinga2007generation}:
\begin{equation}
\frac{1}{\RR_x} = \int_0^\infty \exp(- r_x \tau) g_x(\tau) \dtau.
\end{equation}
We can approximate this $r$--$\RR$ relationship by assuming that the generation-interval distribution is Gamma-distributed, and summarizing it using the mean generation interval $\bar{G}_x$ and the squared coefficient of variation $\kappa_x$ \citep{park2019practical}:
\begin{equation}
\RR_x \approx (1 + \kappa_x r_x \bar{G}_x)^{1/\kappa_x}.
\end{equation}
Various Gamma-generation-interval assumptions have been widely used in epidemic modeling, including for models of SARS-CoV-2 \citep{doi:10.1098/rsif.2020.0144}.
The Gamma-generation-interval assumption includes as a special case models that assume exponentially distributed generation intervals (when $\kappa=1$), corresponding to the SIR model \citep{anderson1991infectious}.
We note that when the infectious periods are Gamma distributed---another standard assumption in epidemic modeling---the resulting generation interval does not necessarily follow the Gamma distribution;
in particular, the mean generation interval can be shorter than the mean infectious period because transmission occurs throughout the infectious period until recovery (see \cite{roberts2007model} for detailed discussion).

We use this framework to investigate how inferences about strength and speed of the variant depend on our assumptions about the underlying generation-interval distributions.
Here, we focus on differences in mean generation intervals, assuming that both the squared coefficient of variation and distribution shape of the variant are otherwise the same as for the wild type: $\kappa_{\mathrm{wt}} = \kappa_{\mathrm{var}} = \kappa$.
Changes in shape can be important \citep{miller2010epidemics,svensson2015influence}, but we do not investigate them here. 
We do note that we expect a distribution with higher coefficient of variation to allow for more early transmission, and thus to have qualitatively similar effects to a distribution with a shorter mean \citep{park2019practical}.

@@

While epidemic speed $r_x$ can often be estimated directly from incidence of infection during the exponential growth period \citep{mills2004transmissibility,nishiura2009transmission,ma2014estimating},
studies of new SARS-CoV-2 variants have mostly focused on characterizing changes in the \emph{proportion} of a new variant \citep{switzerland2021variant, davies2021estimated, di2021impact, graham2021changes, leung2021early, volz2021transmission,zhao2021}.
Focusing on proportions is advantageous, because changes in proportions are less sensitive to changes in testing and to other transient effects that would affect variants and wild type viruses similarly.
When incidence is changing exponentially ($i_x(t) = i_x(t_0) \exp(r_x t)$), the proportion of the new variant $p(t)$ follows a logistic growth curve \citep{switzerland2021variant,davies2021estimated}:
\begin{align}
p(t) &= \frac{\iv(t_0) \exp(\rv t)}{\iw(t_0) \exp(\rw t) + \iv(t_0) \exp(\rv t)},
\\ &= \frac{1}{1 + \left(\iw(t_0)/\iv(t_0)\right) \exp(-\delta t)},
\end{align}
where the relative speed of the variant $\delta$ can be estimated as the slope of the log odds of $p$  vs.~time.
When more than two strains are co-circulating, the picture is more complicated \citep{campbell2021increased}; we focus here on comparing two strains at a time.

\begin{figure}[!t]
\includegraphics[width=\textwidth]{relstrength.pdf}
\caption{
\textbf{Relative strength of the new variant assuming a fixed speed advantage $\delta$ under five epidemiological conditions.}
The relative strength of the new variant $\rho$ (shown on a log scale) conditional on the speed of the wild type $\rw$; the ratio between the mean generation interval of the new variant $\Gv$ and that of the wild type $\Gw$; and the squared coefficient of variation in generation intervals $\kappa$.
The relative strength of the new variant $\rho$ is calculated using $\delta=0.1\pday$, $\Gw = 5\days$, and $\kappa = 1/5$.
Assumed values of $\rw$ and $\rv$ are shown in the top right corners of each panel.
}
\label{fig:relstrength}
\end{figure}

@@

We start by taking the mean generation interval of the wild type to be $\Gw = 5\days$ \citep{ferretti2020quantifying} and the squared coefficient of variation of generation intervals to be $\kappa=0.2$ \citep{ferretti2020quantifying} for both the variant and the wild type.
As noted above, we assume throughout that the variant and the wild type can be approximated with gamma distributions with equal $\kappa$, and only consider differences in the mean.
We evaluate the estimates of relative strength $\rho$ across a wide range of $\kappa$ from 0 (fixed-length generation intervals) to 1 (exponential distributions).
To further explore how inference depends on underlying epidemiological conditions, we consider five scenarios: (1) $\rw < \rv < 0$, (2) $\rw < \rv = 0$, (3) $\rw < 0 < \rv$, (4) $0 = \rw < \rv$, and (5) $0 < \rw < \rv$.
In all of these cases, we assume $\delta = \rv - \rw = 0.1\pday$ based on the observed relative growth rate of the Alpha variant in the UK \citep{davies2021estimated}.
